---
title: "R Project"
author: "Erin Robinson"
date: "8/5/2021"
output:
  md_document:
    variant: markdown_github
    preserve_yaml: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Research Question

My research question is 'Can I reproduce the findings in the Executive Summary of the [State of Open At University of Colorado Baseline Report?](https://scholar.colorado.edu/concern/articles/mw22v6063)'

The findings were:

-   *10.20% of all articles published in 2017 by CU Boulder authors were published in full OA journals;*

-   *From 2013 to 2017, the Libraries OA Fund funded author fees totalling \$163,522.40 for 116 journal articles published by CU Boulder authors in full OA journals;*

-   *As of April 2018, there were 7,732 OA items in the CU Scholar institutional repository run by the Libraries, and these items were downloaded a total of 477,695 times in 2017 alone;*

-   *Published data sets became a reportable scholarly activity in the annual Faculty Report of Professional Activities in 2014, and faculty reported 33 data sets in 2017 (up from 12 in 2014) with 16 of these citations including Digital Object Identifiers (DOIs).*

I used data available from CUScholar\
Johnson, A., Caillet, R., Cantrell, M. H., & Mika, K. (2018). Data from 'State of Open at the University of Colorado Boulder' Report \[Data set\]. University Libraries, University of Colorado Boulder. <https://doi.org/10.25810/ZNB1-3K90>

## Finding 1

-   10.20% of all articles published in 2017 by CU Boulder authors were published in full OA journals;

```{r, echo = TRUE, message = FALSE}
# call Tidyverse library  
library(tidyverse)

# Read in CUBoulderOAArticles2017.csv
CUBoulderOAArticles2017 <- read.csv('data/StateOfOpenCUBoulderData/CUBoulderOAArticles2017.csv')

# Each row is an Open Access article. Use nrow to count number of rows and assign to variable totaloa2017
totaloa2017 <- nrow(CUBoulderOAArticles2017)

# Write out results for total number of OA articles
cat("From the data provided, it is possible to get the total number of open access articles ->", totaloa2017,".")

# Calculate percentage based on average annually number reported in report
perctotal0a2017 <- totaloa2017/5000*100

# Write out results for percentage of total number of OA articles compared to total number of papers 
cat(" The report indicated that CU Boulder publishes around 5000 papers annually. From that statement the percent of OA papers ->", perctotal0a2017, "%. ")

```

No data is provided for the overall number of publication from CU-Boulder for 2017. There is a reference to CU publishing around 5000 papers annually and to CUBE in the text.

Note: It is possible to search the [DataCite Commons for CU-Boulder ROR, organizational identifiers, for 2017](https://commons.datacite.org/ror.org/02ttsq026?rorid=02ttsq026&published=2017). Here we find that there are only 52 works that utilize the ROR that are available. If all research works utilized CU's ROR, we would be able to identify total number of papers published affiliated with CU and then the OA papers/All papers would find % of work published as OA.

## Finding 2

-   *From 2013 to 2017, the Libraries OA Fund funded author fees totaling \$163,522.40 for 116 journal articles published by CU Boulder authors in full OA journals;*

```{r, echo = TRUE, message = FALSE}
# call Tidyverse library  
library(tidyverse)

# Read in CUBoulderOAFund2013_2017.csv
CUBoulderOAFund2013_2017 <- data.frame(read.csv('data/StateOfOpenCUBoulderData/CUBoulderOAFund2013_2017.csv'))

# Explore data by printing a summary of the spreadsheet. This tells us there are six variables, 116 rows and all variables are character strings  
summary(CUBoulderOAFund2013_2017)

```

```{r}
# For Finding 1 we need to sum the column Paid.Amount 
# First we must convert the column from character to numeric class 

# Special characters need to be removed. Remove the $ and , from the Paid.Amount column and assign new variable, Paid.Amountst  
Paid.Amountst <- stringr::str_remove_all(CUBoulderOAFund2013_2017$Paid.Amount,"[$,]")

# Convert the Paid.Amountst character string into a number using as.numeric and assign to a new variable Paid.AmountNum 
Paid.AmountNum <- as.numeric(Paid.Amountst)

# Sum the entire Paid.AmountNum column. There are some fields, so we set na.rm to True. Assign the sum to a new variable called Total.Paid
Total.Paid <- sum(Paid.AmountNum, na.rm = TRUE)

# Print Total.Paid
cat("$",Total.Paid," equals finding 2 From 2013 to 2017, the Libraries OA Fund funded author fees totaling $163,522.40.")


```

```{r}
# Each row represents a journal article, so counting the rows with nrow should give the total number of journal articles 
totaloa2013_2017 <- nrow(CUBoulderOAFund2013_2017)
cat(totaloa2013_2017, "matches the 116 journal articles published by CU Boulder authors in full OA journals from finding 2.")
```

## Finding 3

-   *As of April 2018, there were 7,732 OA items in the CU Scholar institutional repository run by the Libraries, and these items were downloaded a total of 477,695 times in 2017 alone;*

The data shared only included journal publications and data. There was no data included to support this finding.

## Finding 4

-   *Published datasets became a reportable scholarly activity in the annual Faculty Report of Professional Activities in 2014, and faculty reported 33 data sets in 2017 (up from 12 in 2014) with 16 of these citations including Digital Object Identifiers (DOIs).*

```{r, echo = TRUE, message = FALSE}
library(tidyverse)
# Read in CUBoulderOAFund2013_2017 and review the data. There are 7 variables and 87 rows. All variables are character strings 
data <- read.csv('data/StateOfOpenCUBoulderData/CUBoulderPublishedData2014_2017.csv')
colnames(data)
summary(data)

```

```{r}

#Identify the number of datasets in 2014 by filtering data for Year equal to 2014, counting the rows with nrow which equal number of datasets in 2014. Assign the number to a variable data14
data14 <- nrow(filter(data, Year == '2014'))

#Check data14 is correct
#data14

#Repeat process for identifying the number of datasets in 2017 as above for 2014. 
data17 <- nrow(filter(data, Year == '2017'))

#Check data17 is correct
#data17

#Number of datasets with a DOI in 2017 is found by filtering the table to just datasets from 2017 and assigning that to a variable, doi17a
doi17a <- filter(data, Year == '2017')

#doi17a is then filtered again for the variable DOI...yes.or.no equal to yes and nrow is used to count the resulting filtered rows. 
doi17b <-nrow(filter(doi17a, DOI...yes.or.no =='Yes'))

#check doi17b is correct
#doi17b

cat("Print results for Finding 4 -> Published datasets became a reportable scholarly activity in the annual Faculty Report of Professional Activities in 2014, and faculty reported", data17, "data sets in 2017 (up from", data14, "in 2014) with", doi17b," of these citations including Digital Object Identifiers (DOIs).")
```

# 
